---
title: CV
name: Josh M.
surname: London
position: "Wildlife Biologist"
address: "Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA"
email: "josh.london@noaa.gov"
twitter: jomilo75
github: jmlondon
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::markdowncv:
    theme: davewhipp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

## Personal Statement
Wildlife biologist responsible for collaborative design, implementation and
management of research and analysis related to seal ecology and conservation in
Alaska. Species of focus include harbor seals and the ice-associated seals
(ribbon, spotted, bearded, and ringed seal). Research areas of interest include
modeling of marine mammal movement and behavior from deployed tags, survey
design and abundance estimation from aerial surveys, haul-out and dive behavior
and application of new technologies to improve ecological research. Expertise in
R and literate programming approaches (e.g. Rmarkdown) and am increasingly
focussed on open science, reproducibility in ecological analysis, and serve as
an Openscapes (https://www.openscapes.org/) advocate and mentor within AFSC and
NMFS.

## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "B.S.", "1993-1997", "College of Forest Resrouces, Univ. of Washington", "Seattle, WA, USA",
  "Ph.D.", "1998-2006", "School of Aquatic & Fishery Sciences, Univ. of Washington", "Seattle, WA, USA"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

## Dissertation

```{r}
tribble(
  ~ Title, ~ With,  ~Where, ~ When, 
  "Harbor Seals in Hood Canal: Predators and Prey", "Univ. of Washington", "Title:", "1998-2006"
) %>% 
  detailed_entries(what = Title, with = With, when = When, where = Where)
```

Research focused on estimation of predation of endangered summer chum salmon by
harbor seals in Hood Canal, Washington. Additional research efforts to estimate
consumption of harbor seals by Bigg's killer whales. Also expanded on early
techniques for extracting harbor seal DNA from fecal samples to determine sex of
seal for insights into sex-specific foraging ecology.


## Employment

```{r}
tribble(
  ~ Position, ~ Year, ~ Institution, ~ Where,
  "Wildlife Biologist","2006-Present", "Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries", "Position:"
) %>% 
  detailed_entries(
    Position,
    Year, 
    Institution,
    Where
  )
```

## R Packages

```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "crawl","2008-present","https://github.com/NMML/crawl"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _co-lead developer_

The Correlated RAndom Walk Library of R functions was designed for fitting
continuous-time correlated random walk (CTCRW) models with time indexed
covariates. The model is fit using the Kalman-Filter on a state space version of
the continuous-time stochastic movement process.

```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "pathroutr","2021-present","https://github.com/jmlondon/pathroutr"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

The goal of {pathroutr} is to provide functions for re-routing paths that cross
land around barrier polygons. The use case in mind is movement paths derived
from location estimates of marine animals. Due to error associated with these
locations it is not uncommon for these tracks to cross land. The {pathroutr}
package aims to provide a pragmatic and fast solution for re-routing these paths
around land and along an efficient path. It is important to note any re-routed
paths are not statistical models and may have unintended consequences and biases
associated with them that depends on the study animal, study area, and error
nature of the original data. Re-routed paths may also exceed the speed or
physiological capabilities of the study species.


```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "wcUtils","2014-present","https://github.com/jmlondon/wcUtils"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

A package for import/export of Wildlife Computers tag data. {wcUtils} provides
functionality for working with data from select (SPLASH, SPOT) Wildlife
Computers satellite telemetry tags. The package relies on data files produced by
the Wildlife Computers DAP program or downloaded from the Wildlife Computers
Data Portal.


```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "tidextractr","2017-present","https://github.com/jmlondon/tidextractr"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

This package provides a basic interface with the XTide library for estimation of
tidal covariates. The package was developed to support analysis of harbor seal
aerial survey and haul-out behavior in Alaska. However, use is not limited to
this situation. The package does not include the XTide library and, thus, the
library and command line tool must be installed. A vignette will provide some
guidance for installation. Users may wish to investigate the rtide package as an
alternative that includes native support for some of the station locations.
Special thanks and attribution to David Flater for development and support of
the XTide library (http://www.flaterco.com/xtide/xtide.html).

## Publications   
Peer Reviewed

```{r}

bibliography_entries("peer_pubs.bib") 

```

## Publications   
In-Press

```{r}

# bibliography_entries("inpress_pubs.bib")

```

## Publications   
Pre-Print

```{r}

# bibliography_entries("preprint_pubs.bib")

```



