---
title: CV
name: Josh M.
surname: London, Ph. D.
position: "Wildlife Biologist"
address: "Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA"
email: "josh.london@noaa.gov"
twitter: jomilo75
github: jmlondon
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::markdowncv:
    theme: davewhipp
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

## Personal Statement
My career as a wildlife biologist has focused on the ecology and
conservation of seals in Alaska with an emphasis on potential impacts of
climate changes on abundance, trend, and movement ecology. Species of focus
include harbor seals and the ice-associated seals (ribbon, spotted, bearded, and
ringed seals). Research areas of interest include modeling of marine mammal
movement and behavior from deployed bio-loggers, survey design and abundance estimation
from aerial surveys, haul-out and dive behavior, and application of new
technologies to improve ecological research. I am an expert in R and literate
programming approaches (e.g. Rmarkdown) and I am increasingly focused on open
science, reproducibility in ecological analysis, and improving scientific
collaboration.

These guiding principles and ideas often inform my actions:

**Learning is a constant** ... All of the programming, quantitative, and data
science skills I employ on a daily basis have been learned since graduate
school. In some cases, those skills were acquired through formal training
opportunities. In most cases, those skills were self taught and developed
because of my passion for continued learning.

**Find the quantitative within the complexity of ecology** ... I am fortunate to
have close collaborations with some of the finest mathematical statisticians in
statistical ecology. Those collaborations have encouraged and allowed me to seek
out quantitative approaches to address what are often complex questions about
ecosystems and seal ecology.

**Awareness of emerging technologies** ... My commitment to a career of
constant learning has also fostered my radar for emerging technologies in realms
outside of ecology (or outside my regular community) that might have potential
for advancing our understanding of ecosystems. From artificial intelligence 
and machine learning, to thermal imaging capabilities, to advanced satellite
communication platforms, to network routing in mapping applications, or
the culture of open source software development, I'm always looking for 
connections and collaborations.

**Ecological science products should be F.A.I.R.**  
I think ecological data and the science products we produce should follow the
FAIR Data Principles (Findable, Accessible, Interoperable, and Reusable). I
think this is even more relevant as a U.S. government scientist. That said, the
funding, time, knowledge and skills, and technical infrastructure for widespread
application of FAIR principles does not exist within our community. I'm
committed to paving the way, in small efforts, and finding opportunities to
advance these practices when possible.

**Scientific coding is an essential skill** ... This is true not only for myself
but also my collaborators, colleagues, and the greater community of ecological
sciences. Whenever possible, I try to extend my knowledge to those around me,
spend time as a mentor, provide tutorials, and advocate for training and
professional development of scientific coding within NOAA Fisheries.

**Meaningful collaboration requires empathy** ... The nature of my work as an
ecologist and as an employee of NOAA Fisheries always requires collaboration.
Those collaborative efforts are rarely successful without a meaningful
commitment to listening, empathy, and working toward common goals.



## Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "B.S.", "1993-1997", "College of Forest Resrouces, Univ. of Washington", "Seattle, WA, USA",
  "Ph.D.", "1998-2006", "School of Aquatic & Fishery Sciences, Univ. of Washington", "Seattle, WA, USA"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

```{r}
tribble(
  ~ Title, ~ With,  ~Where, 
  "Harbor Seals in Hood Canal: Predators and Prey", "Univ. of Washington", "Title:"
) %>% 
  detailed_entries(what = Title, with = With, where = Where, when = " ")
```

Research focused on estimation of predation of endangered summer chum salmon by
harbor seals in Hood Canal, Washington. Additional research efforts to estimate
consumption of harbor seals by Bigg's killer whales. Also expanded on early
techniques for extracting harbor seal DNA from fecal samples to determine sex of
seal for insights into sex-specific foraging ecology.


## Employment

```{r}
tribble(
  ~ Position, ~ Year, ~ Institution, ~ Where,
  "Wildlife Biologist","2006-Present", "Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries", "Current Position:"
) %>% 
  detailed_entries(
    Position,
    Year, 
    Institution,
    Where
  )
```

As a wildlife biologist with the Polar Ecosystems Program, I am directly
involved in the scientific research and conservation of seals in U.S/Alaska
waters. This work involves collaborative design of research efforts, 
leadership of field activities, co-development and testing of cutting edge
technologies related to aerial surveys and bio-logging, communication of
research results to the broader scientific community and Alaskan
communities, support of the NOAA Fisheries mission of conservation and 
management of protected species, and budget and administrative planning.

**NOAA Fisheries/AFSC Committees & Initiatives**

- RStudio Shiny/Connect Enterprise Service Initiative & Governance Board
- AFSC & NMFS Openscapes Initiatives and Champions Cohorts
- Climate Fisheries Initiative, Blue Team
- NMFS R User-Group Coordination Team

**NOAA Competitive Grants Awarded**

- NMFS Protected Species Toolbox Initiative (co-investigator)
- OAR UAS RFP: Monitoring body condition of seals in Alaska using small UAS (co-author)
- AFSC Regional Work Plan: Trophic roles of ice seals in the Bering and Arctic Marine Ecosystems (co-author)

## Leadership Activies

```{r}
tribble(
  ~ Activity, ~ Year, ~ Location, ~Role, ~Desc,
  "7th International Bio-logging Symposium", "2021", "Virtually from Hawaii", "Co-Organizer & Science Program Chair",
  "575+ attendees from 49 countries and 6 continents"
) %>% 
  detailed_entries(Activity, Year, Location, Role, why = Desc)
```


## Select Committees

```{r}
tribble(
  ~ Committee, ~ Year, ~ Institution, ~ Country, ~ Role,
  "Special Committee on Seals (SCOS)", "2022-2027", "Natural Environment Research Council",
  "United Kingdom","committee chair",
  "Special Committee on Seals (SCOS)", "2019-2021", "Natural Environment Research Council",
  "United Kingdom","scientific advisor",
  "National Marine Mammal Peer Review Committee (NMMPRC) on Northwest Atlantic grey seals",
  "2021", "Fisheries & Oceans Canada", "Canada","external reviewer",
  "US Animal Telemetry Network (ATN) Task Team", "2014-2016", "NOAA/IOOS", "United States",
  "task team member and implementation plan co-author"
) %>% 
  detailed_entries(Committee, Year, Institution, Country, Role)
```

## Awards

```{r}
tribble(
  ~ Award, ~ Year, ~ Institution,
  "Bronze Award", "2009", "NOAA"
) %>% 
  brief_entries(Award, Year, Institution)
```

Group award for successfully planning and implementing the first autonomous
launch and retrieval of an Unmanned Aircraft System on-board a NOAA research
vessel.

## R Packages

```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "**crawl**","2008-present","https://github.com/NMML/crawl"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _co-lead developer_

The Correlated RAndom Walk Library of R functions was designed for fitting
continuous-time correlated random walk (CTCRW) models with time indexed
covariates. The model is fit using the Kalman-Filter on a state space version of
the continuous-time stochastic movement process.

```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "**pathroutr**","2021-present","https://github.com/jmlondon/pathroutr"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

The goal of {pathroutr} is to provide functions for re-routing paths that cross
land around barrier polygons. The use case in mind is movement paths derived
from location estimates of marine animals. Due to error associated with these
locations it is not uncommon for these tracks to cross land. The {pathroutr}
package aims to provide a pragmatic and fast solution for re-routing these paths
around land and along an efficient path. It is important to note any re-routed
paths are not statistical models and may have unintended consequences and biases
associated with them that depends on the study animal, study area, and error
nature of the original data. Re-routed paths may also exceed the speed or
physiological capabilities of the study species.


```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "**wcUtils**","2014-present","https://github.com/jmlondon/wcUtils"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

A package for import/export of Wildlife Computers tag data. {wcUtils} provides
functionality for working with data from select (SPLASH, SPOT) Wildlife
Computers satellite telemetry tags. The package relies on data files produced by
the Wildlife Computers DAP program or downloaded from the Wildlife Computers
Data Portal.


```{r}
tribble(
  ~ PkgName, ~ReleaseDate, ~URL,
  "**tidextractr**","2017-present","https://github.com/jmlondon/tidextractr"
) %>% 
  brief_entries(PkgName, ReleaseDate, URL)
```

- _lead developer_

This package provides a basic interface with the XTide library for estimation of
tidal covariates. The package was developed to support analysis of harbor seal
aerial survey and haul-out behavior in Alaska. However, use is not limited to
this situation. The package does not include the XTide library and, thus, the
library and command line tool must be installed. A vignette will provide some
guidance for installation. Users may wish to investigate the rtide package as an
alternative that includes native support for some of the station locations.
Special thanks and attribution to David Flater for development and support of
the XTide library (http://www.flaterco.com/xtide/xtide.html).

## Publications   
Peer Reviewed

```{r}

bibliography_entries("peer_pubs.bib") %>% 
  mutate(issue_date = as.Date(issued)) %>% 
  arrange(desc(issue_date))

```

## Publications   
Technical Reports

```{r}

bibliography_entries("tech_reports.bib")%>% 
  mutate(issue_date = as.Date(issued)) %>% 
  arrange(desc(issue_date))

```

## Publications   
Pre-Print

```{r}

# bibliography_entries("preprint_pubs.bib")

```

## Presentations
Oral and Poster

```{r}
bibliography_entries("presentations.bib")%>% 
  mutate(issue_date = as.Date(issued)) %>% 
  arrange(desc(issue_date))
```


